### 2011-12-25 doc/theses/mikula/proposed_algorithm.tex
Master thesis: In progress.

A simple algorithm is presented in Algorithm \ref{ALG_determination_of_function_return_types} \todo{Q: uvadzat aj taketo jednoduche funkcie? A: Obecně by mělo být popsáno vše. Ale pokud se nějaká jednoduchá věc např. opakuje, tak stačí napsat, že je algoritmus podobný jako v předchozím případě nebo tak něco.}. It uses \texttt{getFunctionDeclarationNodes()} function, which returns a list of all function declaration nodes in the syntax three. Actually, it does not have to search the whole syntax tree as the function declaration nodes can be present only in the query prolog section.

@@

\begin{algorithm}
\caption{Determination of Function Return Types}
\label{ALG_determination_of_function_return_types}
\begin{algorithmic}[1]
\ENSURE An associative array of function names with their types.

\STATE $functionTypes$ = empty associative array
\FORALL{$functionDeclarationNode \in getFunctionDeclarationNodes()$}
    \STATE $functionName = functionDeclarationNode.getFunctionName()$
    \STATE $functionTypes[functionName] = functionDeclarationNode.getReturnType()$
\ENDFOR
\RETURN $functionTypes$
\end{algorithmic}
\end{algorithm}

For the rest of the thesis, we assume function \todo{Většinou se na začátek sekce s algoritmem dává seznam používaných funkcí a jejich význam.} \texttt{getFunctionReturnType()} which takes a function name and returns the return type of the function if the function is either built-in or it is recorded in the associative array. Otherwise, \texttt{null} is returned.

Later phases of the algorithm may need to process the function declarations. Therefore, we also store references to the entire function declaration nodes.

@@

\STATE $type = null$
\IF{$isA(exprNode, LiteralNode)$}
     \STATE $type =$ ((LiteralNode)exprNode).type, exactly-once cardinality
\ELSIF{$isA(exprNode, FunctionCallNode)$}
     \STATE $type = getFunctionReturnType(((FunctionCallNode)exprNode).functionName)$
\ELSIF{$isA(exprNode, VarRefNode)$}
     \STATE $type = getVariableType(contextVarTypes, ((VarRefNode)exprNode).variableName)$
\ELSIF{$isA(exprNode, PathExprNode)$}
     \STATE $path = ((PathExprNode)exprNode).path$
     \IF{$path$ selects an element set}
         \STATE $type =$ element type, zero-or-more cardianlity, $path$ included
     \ELSIF{$path$ selects an attribute set}
         \STATE $type =$ attribute type, zero-or-more cardianlity, $path$ included
     \ELSE
         \STATE $type = null$
     \ENDIF
\ELSIF{$isA(exprNode, OperatorNode)$}
     \STATE $type = determineOperatorType(exprNode, contextVarTypes)$
\ELSIF{$isA(exprNode, FLWORExprNode)$}
     \STATE $returnClauseNode = ((FLWORExprNode)exprNode).returnClauseNode$
     \STATE $type = returnClauseNode.type$
\ENDIF

>>>>

\begin{algorithm}
\caption{Retrieval of Function Declaration Nodes $getFunctionDeclarationNodes(syntaxTree)$}
\label{ALG_retrieval_of_function_declaration_nodes}
\begin{algorithmic}[1]
\REQUIRE $syntaxTree$: A reference to the root node of a syntax tree.
\ENSURE A list of syntax tree nodes representing function declarations.

\STATE $prologNode = null$
\FORALL{$moduleChildNode \in syntaxTree.getChildren()$}
    \IF{$is(moduleChildNode, PrologNode)$}
        \STATE $prologNode = moduleChildNode$
    \ENDIF
\ENDFOR

\STATE $functionDeclarationNodes =$ an empty list
\IF{$prologNode \neq null$}
    \FORALL{$prologChildNode \in prologNode.getChildren()$}
        \IF{$is(prologChildNode, FunctionDeclNode)$}
            \STATE add $prologChildNode$ to $functionDeclarationNodes$
        \ENDIF
    \ENDFOR
\ENDIF

\RETURN $functionDeclarationNodes$
\end{algorithmic}
\end{algorithm}

A simple algorithm is presented in Algorithm \ref{ALG_determination_of_function_return_types} \todo{Q: uvadzat aj taketo jednoduche funkcie? A: Obecně by mělo být popsáno vše. Ale pokud se nějaká jednoduchá věc např. opakuje, tak stačí napsat, že je algoritmus podobný jako v předchozím případě nebo tak něco.}. It uses \texttt{getFunctionDeclarationNodes()} function, defined in Algorithm \ref{ALG_retrieval_of_function_declaration_nodes}, which returns a list of all function declaration nodes in the syntax three. Actually, it does not have to search the whole syntax tree as the function declaration nodes can be present only in the query prolog section.

@@

\begin{algorithm}
\caption{Determination of Function Return Types}
\label{ALG_determination_of_function_return_types}
\begin{algorithmic}[1]
\REQUIRE{\ \\
    $getFunctionDeclarationNodes()$: Function defined in Algorithm \ref{ALG_retrieval_of_function_declaration_nodes}. \\
    $syntaxTree$: A reference to the root node of a syntax tree.
}
\ENSURE An associative array of function names with their types and references to the their declaration nodes.

\STATE $functionArray$ = an empty associative array
\FORALL{$functionDeclarationNode \in getFunctionDeclarationNodes(syntaxTree)$}
    \STATE $functionName = functionDeclarationNode.funcName$
    \STATE $typeNode = functionDeclarationNode.getChild(TypeNode)$
    \STATE $functionArray[functionName] = (PROCESS\_TYPE(typeNode), functionDeclarationNode)$
\ENDFOR
\RETURN $functionArray$
\end{algorithmic}
\end{algorithm}

\todo[inline]{Definovat a premenovat PROCESS\_TYPE() v Determination of Function Return Types. V skratke robi to, ze instanciu TypeNode prevediu na nasu reprezentaciu typu.}

For the rest of the thesis, we assume function \todo{Většinou se na začátek sekce s algoritmem dává seznam používaných funkcí a jejich význam.} $getFunctionReturnType()$ which takes a function name and returns the return type of the function if the function is either built-in or it is recorded in the associative array. Otherwise, \texttt{null} is returned.

Later phases of the algorithm may need to process the function declarations. Therefore, we also store references to the entire function declaration nodes, accessible using function $getFunctionDeclNode()$.

@@

\STATE $type = null$
\IF{$is(exprNode, LiteralNode)$}
     \STATE $type = exprNode.type$, exactly-once cardinality
\ELSIF{$is(exprNode, FunctionCallNode)$}
     \STATE $type = getFunctionReturnType(exprNode.functionName)$
\ELSIF{$is(exprNode, VarRefNode)$}
     \STATE $type = getVariableType(contextVarTypes, exprNode.variableName)$
\ELSIF{$is(exprNode, PathExprNode)$}
     \STATE $path = exprNode.path$
     \IF{$path$ selects an element set}
         \STATE $type =$ element type, zero-or-more cardinality, $path$ included
     \ELSIF{$path$ selects an attribute set}
         \STATE $type =$ attribute type, zero-or-more cardinality, $path$ included
     \ELSE
         \STATE $type = null$
     \ENDIF
\ELSIF{$is(exprNode, OperatorNode)$}
     \STATE $type = determineOperatorType(exprNode, contextVarTypes)$
\ELSIF{$is(exprNode, FLWORExprNode)$}
     \STATE $returnClauseNode = exprNode.returnClauseNode$
     \STATE $type = returnClauseNode.type$
\ENDIF
